Multi-level Selective Deduplication for VM Snapshots in Cloud Storage

### 2012-03-09 vm_snapshot_sim/2012abs.tex
updated.

\begin{abstract}
%Virtualization has became the engine behind many cloud computing platforms.
In a virtualized cloud computing environment, frequent  snapshot backup of virtual disks improves
hosting  reliability while  storage demand of such operations is huge.
A dirtybit-based technique can identify unmodified data between versions 
and full deduplication with fingerprint comparison  can remove more redundant content
while it requires more computing resource.
%with  for similarity comparison and   reliability handling.
%Current snapshot deduplication is mainly done through copy-on-write 
%on fixed-size disk blocks. Such solutions cannot handle the
% cross VM data duplication because VMs do not share any data. 
%In addition, storing VM images and their snapshots
%in the same storage engine reduce the underline design flexibility because 
%these two kinds of data have distinct access requirements.
%In this paper, 
%we show that there is a large amount of duplicated data shared amongy virtual machines
%through a production VM data study and thus it is expective to perform cross-machine deduplication. 
This paper presents a multi-level selective deduplication  scheme which
integrates  inner-VM and cross-VM duplicate elimination under
a stringent resource requirement.
% minimal  resource impact to the existing cloud services.  
This scheme uses popular common data to facilitate 
fingerprint comparison while reducing the cost and it
strikes a balance  between local and global deduplication 
to  increase parallelism and  improve reliability. 
% first perform a large scale study in production VM clusters 
%to show that cross VM data duplication is severe due to they have large amount of
%common data. Then our data analysis finds out that the overall data duplication pattern follows the Zipf's law.
%Base on these discoveries, we propose a snapshot storage deduplication scheme using variable-size chunking
%to address the above problem efficiently.
%We eliminate the majority of cross VM data duplication by pre-select
%a small set of frequently seen data blocks to be shared globally, and we also remove
%many cross snapshot duplication by using smaller chunking granuarity and locality.
Experimental results  show the proposed scheme  can achieve high deduplication ratio while using
a  small  amount of cloud resources. 
\end{abstract}

>>>>

\begin{abstract}
%Virtualization has became the engine behind many cloud computing platforms.
In a virtualized cloud computing environment, frequent  snapshot backup of virtual disks improves
hosting  reliability while  storage demand of such operations is huge.
While dirtybit-based technique can identify unmodified data between versions, 
full deduplication with fingerprint comparison  can remove more redundant content
while it requires more computing resources.
%with  for similarity comparison and   reliability handling.
%Current snapshot deduplication is mainly done through copy-on-write 
%on fixed-size disk blocks. Such solutions cannot handle the
% cross VM data duplication because VMs do not share any data. 
%In addition, storing VM images and their snapshots
%in the same storage engine reduce the underline design flexibility because 
%these two kinds of data have distinct access requirements.
%In this paper, 
%we show that there is a large amount of duplicated data shared amongy virtual machines
%through a production VM data study and thus it is expective to perform cross-machine deduplication. 
This paper presents a multi-level selective deduplication  scheme which
integrates  inner-VM and cross-VM duplicate elimination under
a stringent resource requirement.
% minimal  resource impact to the existing cloud services.  
This scheme uses popular common data to facilitate 
fingerprint comparison while reducing the cost and it
strikes a balance  between local and global deduplication 
to  increase parallelism and  improve reliability. 
% first perform a large scale study in production VM clusters 
%to show that cross VM data duplication is severe due to they have large amount of
%common data. Then our data analysis finds out that the overall data duplication pattern follows the Zipf's law.
%Base on these discoveries, we propose a snapshot storage deduplication scheme using variable-size chunking
%to address the above problem efficiently.
%We eliminate the majority of cross VM data duplication by pre-select
%a small set of frequently seen data blocks to be shared globally, and we also remove
%many cross snapshot duplication by using smaller chunking granuarity and locality.
Experimental results  show the proposed scheme  can achieve high deduplication ratio while using
a  small  amount of cloud resources. 
\end{abstract}

### 2012-03-09 vm_snapshot_sim/2012background.tex
updated.

VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%Even though the snapshots are saved incrementally,
%when deleting a snapshot, only the data not needed for any other snapshot is removed.
%So regardless of which prior snapshots have been deleted, all active snapshots will contain
%all the information needed to restore the virtual disk.
%One widely used snapshot strategy is copy-on-write (CoW).
%The earlier use of COW is to improve memory usage and reduce copying overhead in process 
%management~\cite{OSbook,Waldspurger2002} and it can be extended for snapshot 
%storage management~\cite{vmware_kb1015180}. 
%C. Waldspurger. Memory Resource Management in VMware ESX Server in Proceedings of the 5th Symposium 
%on Operating Systems Design and Implementation, 2002
%Upon VM image storage system receives a save snapshot request,
%it freezes the state of that image file, then all consequent write request will result in the write region being copied
%to an incremental snapshot data file. 
%Such a strategy exploits version difference between consecutive snapshots
%of the same image.
The main weakness
is that it does not reveal content redundancy among data blocks from different files or
different VMs.
%has several disadvantages:
%first, CoW may affect the general I/O performance due to defered data coping. 
%Second, CoW does not seperate backup data and runtime image data,
%which have distinct access requirements: runtime image data is directly used by the running VM, 
%thus need high throughput, and is very sensitive to latency,
%such data must be served with hig cost hardware, but backup data generally only need fair aggregate throughput, 
%is not sensitive to latency, thus can be stored in secondary storage devices.
%Finally, VM snapshots contain tremendous amount of data duplication, which is nearly impossible to tackle 
%if these two kinds of data are tightly coupled together.

@@

Data deduplication has been applied as  a post-processing operation for offline removal
of redundant content in the file system data management once backup data is written in a temporary stage area.
Dynamic inline deduplication is an effort to improve storage efficiency  
by detecting and removing  duplicates before data is written to the backup storage,
thus reduces extra disk space required to hold temporary data. 
This approach requires a fingerprint  comparison of the new content chunks with
previously stored fingerprints. 
}
As data grows to be big, content chunk lookup for fingerprint comparison from disk storage
becomes too slow to be scalable.
%and searching such disk
%However, unless some form of locality or similarity is exploited, inline, chunk-based deduplication, 
%when done at a large scale faces what has been termed the disk bottleneck problem: to facilitate fast chunk ID lookup, 
%a single index containing the chunk IDs of all the backed up chunks must be maintained. 
%As the backed up data grows, the index overflows the amount of RAM available and must be paged to disk. 
%Without locality, the index cannot be cached effectively, and it is common for nearly 
%every index access to require a random disk access. This disk bottleneck severely limits deduplication throughput.
Several techniques have been proposed to speedup searching of duplicate 
content~\cite{bottleneck08,extreme_binning09,sparseindex09}. For example,  
Zhu et al.~\cite{bottleneck08} tackle it 
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be 
accessed together with high probability. It takes significant memory resource for filtering.
NG et al.~\cite{ NGmiddleware2011}  use  
a related filtering technique for integrating deduplication in Linux  file system and the memory
consumed is up to 2GB for a single machine. That is still too big in our context discussed below. 
%Lillibridge et al.~\cite{sparseindex09} break list of chunks 
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is 
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments. 
%These are segments that share many chunk IDs with the incoming segment with high probability.
%Deepavali et al.~\cite{extreme_binning09}  use signature-based file similarity  and group similar files
%into the same physical location (bins) to deduplicate against each other.

Duplicate  search approximation~\cite{extreme_binning09,sparseindex09}  has been proposed 
%in extreme bining and other sparse indexing 
to package similar content in one disk, and duplicate lookup  only searches
for chunks within files which have a similar file-level or segment-level  content fingerprints.
That leads  to a smaller amount of memory usage for storing meta data in signature
lookup with a tradeoff of the reduced recall ratio.

@@

We discuss the characteristics and 
main requirements for VM snapshot backup in a cloud environment.
which are different from a traditional data backup. 
% That arises mainly in Alibaba's Aliyun cloud service.
%concerns
%Base on Aliyun's production environment, the snapshot backup job has to satisfy 
\begin{enumerate}
\item {\em Cost consciousness.}
There are tens of thousands of VMs running on a large-scale cloud cluster. 
The amount of VM data to backup is huge and cost of complete backup can be prohibitively high without
deduplication and compression support.  Thus cost minimization is required to deliver a deplorable solution.

On the other hand, the computing resource allocated for backup is often very small.
This is because snapshot data is not retrieved frequently by users and the allocation of  cluster resource
is typically prioritized towards active services such as online web services and map-reduce jobs.
Thus  it is not easy to accept any backup behavior that affects VMs performance or stability of active users.  
At Aliyun, it is required that while CPU and disk usage should be  small or modest during backup time,
the memory footprint of deduplication should not exceed 1\% of memory at each node, which is about 480MB.

%snapshot service shall not compete cpu, memory, or I/O bandwidth with VMs. Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
\item {\em Fast backup speed.}
The snapshot  backup  operations need to be completed quickly  every day (e.g. 1-2 hours ). 
Often a cloud service has  few hours of light workload each day (e.g. midnight),  which creates an opportunity  for backup.
But a  longer use of bandwidth and computing resource  for backup can create  noticeable  contention with the existing cloud,
which is not preferred for cloud production system operation. Thus it is desirable that backup for all nodes
can be conducted in parallel and any centralized or  cross-machine communication for
deduplication should not become a bottleneck.
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed is huge. 
%For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, The aggreated amount of data 
 % the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.

% the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.
\item {\em Fault tolerance.}
The addition of a backup service should not decrease the degree of
fault tolerance.  If data failure in  a backup service  affects many users,  such a solution is not desirable.
%when users access snapshots in a recovery process. 
\end{enumerate}

>>>>

VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%Even though the snapshots are saved incrementally,
%when deleting a snapshot, only the data not needed for any other snapshot is removed.
%So regardless of which prior snapshots have been deleted, all active snapshots will contain
%all the information needed to restore the virtual disk.
%One widely used snapshot strategy is copy-on-write (CoW).
%The earlier use of COW is to improve memory usage and reduce copying overhead in process 
%management~\cite{OSbook,Waldspurger2002} and it can be extended for snapshot 
%storage management~\cite{vmware_kb1015180}. 
%C. Waldspurger. Memory Resource Management in VMware ESX Server in Proceedings of the 5th Symposium 
%on Operating Systems Design and Implementation, 2002
%Upon VM image storage system receives a save snapshot request,
%it freezes the state of that image file, then all consequent write request will result in the write region being copied
%to an incremental snapshot data file. 
%Such a strategy exploits version difference between consecutive snapshots
%of the same image.
The main weakness
is that it does not reveal content redundancy among data blocks from different snapshots or
different VMs.
%has several disadvantages:
%first, CoW may affect the general I/O performance due to defered data coping. 
%Second, CoW does not seperate backup data and runtime image data,
%which have distinct access requirements: runtime image data is directly used by the running VM, 
%thus need high throughput, and is very sensitive to latency,
%such data must be served with hig cost hardware, but backup data generally only need fair aggregate throughput, 
%is not sensitive to latency, thus can be stored in secondary storage devices.
%Finally, VM snapshots contain tremendous amount of data duplication, which is nearly impossible to tackle 
%if these two kinds of data are tightly coupled together.

@@

Data deduplication has been applied as  a post-processing operation for offline removal
of redundant content in the file system data management once backup data is written in a temporary stage area.
Dynamic inline deduplication is an effort to improve storage efficiency  
by detecting and removing  duplicates before data is written to the backup storage,
thus reduces extra disk space required to hold temporary data. 
This approach requires a fingerprint  comparison of the new content chunks with
previously stored fingerprints. 
}
As data grows to be big, fingerprint lookup in such schemes
becomes too slow to be scalable.
%and searching such disk
%However, unless some form of locality or similarity is exploited, inline, chunk-based deduplication, 
%when done at a large scale faces what has been termed the disk bottleneck problem: to facilitate fast chunk ID lookup, 
%a single index containing the chunk IDs of all the backed up chunks must be maintained. 
%As the backed up data grows, the index overflows the amount of RAM available and must be paged to disk. 
%Without locality, the index cannot be cached effectively, and it is common for nearly 
%every index access to require a random disk access. This disk bottleneck severely limits deduplication throughput.
Several techniques have been proposed to speedup searching of duplicate 
content. For example,  
Zhu et al.~\cite{bottleneck08} tackle it 
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be 
accessed together with high probability. It takes significant memory resource for filtering and caching.
NG et al.~\cite{ NGmiddleware2011}  use  
a related filtering technique for integrating deduplication in Linux  file system and the memory
consumed is up to 2GB for a single machine. That is still too big in our context discussed below. 
%Lillibridge et al.~\cite{sparseindex09} break list of chunks 
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is 
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments. 
%These are segments that share many chunk IDs with the incoming segment with high probability.
%Deepavali et al.~\cite{extreme_binning09}  use signature-based file similarity  and group similar files
%into the same physical location (bins) to deduplicate against each other.

Duplicate  search approximation~\cite{extreme_binning09,sparseindex09}  has been proposed 
%in extreme bining and other sparse indexing 
to package similar content in one location, and duplicate lookup  only searches
for chunks within files which have a similar file-level or segment-level  content fingerprints.
That leads  to a smaller amount of memory usage for storing meta data in signature
lookup with a tradeoff of the reduced recall ratio.

One weakness the hasn't been addressed by previous solutions is that we want to isolate 
each VM's snapshot backup as much as possible while still enjoy the benefit of deduplication.
In large scale cloud, node faiures happen at daily basis, we don't want a problem at small scale
to affect large amount of VMs due to data sharing.

@@

We discuss the characteristics and 
main requirements for VM snapshot backup in a cloud environment.
which are different from a traditional data backup. 
% That arises mainly in Alibaba's Aliyun cloud service.
%concerns
%Base on Aliyun's production environment, the snapshot backup job has to satisfy 
\begin{enumerate}
\item {\em Cost consciousness.}
There are tens of thousands of VMs running on a large-scale cluster. 
The amount of data to backup is so huge such that 3rd-party solutions are generally not affordable.
On the other hand, the computing resources allocated for snapshot service is very limited
because VM performance has higher priority.  
At Aliyun, it is required that while CPU and disk usage should be small or modest during backup time,
the memory footprint of snapshot service should not exceed 500MB at each node.

%snapshot service shall not compete cpu, memory, or I/O bandwidth with VMs. Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
\item {\em Fast backup speed.}
Often a cloud has a few hours of light workload each day (e.g. midnight),  which creates an small window for automatic backup.
%But a  longer use of bandwidth and computing resource  for backup can create  noticeable  contention with the existing cloud,
%which is not preferred for cloud production system operation. 
Thus it is desirable that backup for all nodes
can be conducted in parallel and any centralized or  cross-machine communication for
deduplication should not become a bottleneck.
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed is huge. 
%For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, The aggreated amount of data 
 % the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.

% the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.
\item {\em Fault tolerance.}
The addition of data deduplication should not decrease the degree of
fault tolerance. It's not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process. 
\end{enumerate}

### 2012-03-09 vm_snapshot_sim/2012intro.tex
updated.

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host 
%tens of thousands of  virtual machines(VM) dynamically created
%for various active users. 
%In the  case of Alibaba cloud service, the sevice is called Aliyu, which
%provides the largest public cloud in based on the open-source Xen technology.
In a virtualized cloud environment such as ones provided by Amazon EC2 and  Alibaba Aliyun,
each instance of a guest operating system runs on a virtual machine, accessing
virtual disks represented as virtual disk image files  in the host operating system.
%virtual disk image files (e.g. .vhd, .vmdk) in the host operating system.
Because these image files are stored as regular files from the external point of view,
backing up VM's data is mainly done by taking snapshots of virtual disk images.

@@

Frequent  backup of VM snapshots increases  the reliability of VM's hosted in a cloud.
For example, Aliyun, the largest cloud service provided by Alibaba in China, 
provides automatic frequent backup of VM images to strengthen the reliability of its service for all users.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
Using a backup service with full deduplication support~\cite{venti02,bottleneck08,extreme_binning09,sparseindex09}
can identify content duplicates among snapshots to remove redundant storage content,  but the weakness is that it
either adds the  extra cost significantly or competes computing resource with the existing cloud services.
Data dependence created by duplicate relationship among snapshots
adds the complexity in fault tolerance management, especially when  VMs can migrate around in the cloud.

Unlike the previous work dealing with general file-level backup and deduplication, our problem is focused on 
virtual disk image backup and even each virtual disk is viewed as a file logically, its size is very large.
On the other hand, we need to support parallel backup of a large number of virtual disks in a cloud everyday. 
One key requirement we face at Alibaba Aliyun is that VM snapshot backup should only use a minimal amount of system
resource so that most of resource is allocated for regular cloud system services or applications.
Thus our objective is to exploit the characteristics of VM snapshot data and
pursue a cost-effective deduplication solution. 
Another goal  is to decentralize VM snapshot backup and  localize  deduplication as much as possible,
which brings the benefits for increased parallelism  and fault isolation.

With these in mind, we  have developed a decentralized multi-level solution to conduct 
segment-level  and block-level  within each VM to localize deduplication when possible.
It then makes global inter-VM deduplication efforts by using a small number of
popular common data blocks.  Our study shows that there are a larger number
of popular common content  blocks shared by many users while they only take
a small amount of resource for global deduplication. 
Leveraging popular blocks makes the fault tolerance management easier while
it can still effectively accomplish a significant percentage of inter-VM  deduplication
compared to a full deduplication algorithm. 
%Restricting the scope of global deduplication reduces
%the inter-component  data dependenc during machine failure.

>>>>

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host 
%tens of thousands of  virtual machines(VM) dynamically created
%for various active users. 
%In the  case of Alibaba cloud service, the sevice is called Aliyu, which
%provides the largest public cloud in based on the open-source Xen technology.
In a virtualized cloud environment such as ones provided by Amazon EC2 and Alibaba Aliyun,
each instance of a guest operating system runs on a virtual machine, accessing
virtual hard disks represented as virtual disk image files in the host operating system.
%virtual disk image files (e.g. .vhd, .vmdk) in the host operating system.
Because these image files are stored as regular files from the external point of view,
backing up VM's data is mainly done by taking snapshots of virtual disk images.

@@

Frequent  backup of VM snapshots increases  the reliability of VM's hosted in a cloud.
For example, Aliyun, the largest cloud service provider by Alibaba in China, 
provides automatic frequent backup of VM images to strengthen the reliability of its service for all users.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
Using a backup service with full deduplication support~\cite{venti02,bottleneck08}
can identify content duplicates among snapshots to remove redundant storage content,  but the weakness is that it
either adds the  extra cost significantly or competes computing resource with the existing cloud services.
In addition, data dependence created by duplicate relationship among snapshots
adds the complexity in fault tolerance management, especially when  VMs can migrate around in the cloud.

Unlike the previous work dealing with general file-level backup and deduplication, our problem is focused on 
virtual disk image backup. Although we treat each virtual disk  as a file logically, its size is very large.
On the other hand, we need to support parallel backup of a large number of virtual disks in a cloud everyday. 
One key requirement we face at Alibaba Aliyun is that VM snapshot backup should only use a minimal amount of system
resources so that most of resources is kept for regular cloud system services or applications.
Thus our objective is to exploit the characteristics of VM snapshot data and
pursue a cost-effective deduplication solution. 
Another goal  is to decentralize VM snapshot backup and  localize  deduplication as much as possible,
which brings the benefits for increased parallelism  and fault isolation.

%,extreme_binning09,sparseindex09
By observations on the VM snapshot data from production cloud, we found snapshot data duplication 
can be easily classified into two categories: \emph{inner-VM} and \emph{cross-VM}. Inner-VM duplication
exists between VM's snapshots, because the majority of data are unchanged during each backup period. 
On the other hand, Cross-VM duplication is mainly due to widely-used software and libraries: 
there are only two type of OSes (Linux and Windows), most users either uses MySQL or MS SQL Server as their database.
As the result, different VMs tend to backup large amount of highly similar data.

With these in mind, we  have developed a decentralized multi-level solution to conduct 
segment-level  and block-level inner-VM  deduplication to localize the deduplication effort when possible.
It then makes cross-VM deduplication by excluding a small number of
popular common data blocks from being backed up. Our study shows that common data blocks
occupy significant amount of storage space while they only take
a small amount of resources to deduplicate.
Seperating deduplication into multi levels effectively accomplish the major space saving goal
compare the global complete deduplication scheme, at the same time it makes
the backup of different VMs to be indepdent for better fault tolerance.

%Restricting the scope of global deduplication reduces
%the inter-component  data dependenc during machine failure.
